\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web

### 2016-09-20 revision2/article.tex
revised from reviewer comments

\documentclass[article, shortnames]{jss}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{booktabs}
\usepackage{fixltx2e}

@@

% citation alias for LUBW institution
\defcitealias{lubw_2016}{LUBW (2016)}

\begin{document}

<<global_options, include=FALSE>>=
library(knitr)
knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
knitr::render_sweave()
knitr::opts_chunk$set(prompt = TRUE,
                      comment = NA,
                      highlight = FALSE,
                      dev.args = list(pointsize = 11),
                      # fig.width = 4.9,
                      # fig.height = 3.675,
                      fig.align = 'center',
                      out.format = 'sweave',
                      tidy = FALSE,
                      fig.path = '')
# 
hook.c <- function(x, options) {
  paste0('\\begin{CodeChunk}\n', x, '\\end{CodeChunk}')
}
hook.i <- function(x, options) {
  paste0(c('\\begin{CodeInput}', knitr:::hilight_source(x, 'latex', options), '\\end{CodeInput}', ''), collapse = '\n')
}
hook.s <- function(x, options) {
  paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
}
knitr::knit_hooks$set(chunk = hook.c,
               source = hook.i,
               output = hook.s,
               message = hook.s,
               warning = hook.s,
               plot = knitr::hook_plot_tex)

options(prompt = "R> ", continue = "+  ", width = 76, useFancyQuotes = FALSE)
@

@@

<<install, eval=FALSE>>=
install.packages("webchem")
library("webchem")
@

<<load, echo=FALSE, eval=TRUE>>=
library(webchem)
library(ggplot2)
library(xtable)
@

@@

<<loadjagst>>=
data("jagst")
head(jagst)
@

@@

<<load_lc50>>=
data("lc50")
head(lc50)
@

@@

<<get_etoxid, message=FALSE, cache=TRUE>>=
subs <- unique(jagst$substance)
ids <- get_etoxid(subs, match = 'best')
head(ids)
@

@@

<<etox_basic, message=FALSE>>=
etox_data <- etox_basic(ids$etoxid)
@

\pkg{webchem} always returns a named list (one entry for each substance) and the available information content can be very voluminous.
Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
<<etox_cas, message=FALSE>>=
etox_cas <- cas(etox_data)
head(etox_cas)
@

@@

In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
<<token, echo=FALSE>>=
# this is a webchem specific token. 
# Please use it only for reproduction / testing.
token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
@

<<pubchem, message = FALSE, warning=FALSE>>=
cids <- get_cid(etox_cas)
pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
pc_smiles <- smiles(pc_data)
@

or ChemSpider
<<chemspider, message=FALSE, warning=FALSE>>=
csids <- get_csid(etox_cas, token = token)
cs_data <- cs_compinfo(csids, token = token)
cs_inchikey <- inchikey(cs_data)
@

Finally, we combine the queried data into one data.frame
<<table>>=
res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
  cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
  stringsAsFactors = FALSE)
@

@@

\begin{table}[ht]
\centering
<<print_table, echo=FALSE, results = 'asis'>>=
tmp <- res[1:6, ]
tmp$smiles <- ifelse(is.na(tmp$smiles), NA, paste0(substr(tmp$smiles, 1, 7), '...'))
tmp$inchikey <- ifelse(is.na(tmp$inchikey), NA, paste0(substr(tmp$inchikey, 1, 7), '...'))
tmp[is.na(tmp)] <- '-'
colnames(tmp) <- c("Name", "CAS", "SMILES","CID", "InChIKey", "CSID")
print(xtable(tmp),
  floating = FALSE,
  hline.after = NULL,
  include.rownames = FALSE,
  add.to.row = list(pos = list(-1,0, 6),
                    command = c('\\toprule\n','\\midrule\n','\\bottomrule\n')),
  sanitize.text.function = function(x){x}
  )
@
\caption{Identifiers for the jagst data sets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
\label{tab:comptable}
\end{table}

@@

<<aw_query, message = FALSE>>=
aw_data <- aw_query(lc50$cas, type = 'cas')
@

@@

<<subactivty, message = FALSE>>=
igroup <- sapply(aw_data, function(y) y$subactivity[1])
igroup[1:3]
@

@@

\begin{figure}[ht]
<<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
              ifelse(grepl('neonicotinoid', igroup), 'Neo-\nnicotinoids', 
                ifelse(grepl('pyrethroid', igroup), 'Pyrethroids', 
                  ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
                         'other'))))
lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", "Organo-\nphosphates", "Neo-\nnicotinoids", "other"))

p <- ggplot(lc50, aes(x = type, y = value)) +
  geom_boxplot(fill = 'grey75') + 
  geom_jitter(width = 0.5) +
  scale_y_log10() +
  labs(y = expression(LC[50~'D.magna, 48h']), x = '') +
  theme_bw() +
  theme(text = element_text(size = 16))
p
@
\caption{Toxicity of different pesticide groups. LC\textsubscript{50} values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
\label{fig:fig2}
\end{figure}

@@

<<qsar, message=FALSE, cache=TRUE, warning=FALSE>>=
pp_data <- pp_query(lc50$cas)
@

@@

<<qsar2, warnings = FALSE>>=
lc50$logp <- sapply(pp_data, function(y) {
  if (length(y) == 1 && is.na(y))
    return(NA)
  y$prop$value[y$prop$variable == 'Log P (octanol-water)']
})
@

@@

\begin{figure}[ht]
<<plot_qsar, echo=FALSE, out.width='0.7\\textwidth', fig.height=6, warning=FALSE>>=
# model
mod <- lm(log10(value) ~ logp, data = lc50)
coefs <- round(coef(mod), 2) # coeficients
rmse <- round(summary(mod)$sigma, 2) # RMSE

# check data types
logp_type <- sapply(pp_data, function(y){
  if (length(y) == 1 && is.na(y))
    return(NA)
  y$prop$type[y$prop$variable == 'Log P (octanol-water)']
})
tabs <- table(logp_type)
sna <- sum(is.na(logp_type))

p <- ggplot(lc50, aes(x = logp, y = log10(value))) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(x = expression(log~P[oct/wat]), y = expression(log~LC[50])) +
  theme_bw()
p
@
\caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
Blue line indicates the regression model ($\mathrm{log~LC}_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]}~\mathrm{log} P$, RMSE = \Sexpr{rmse}).}
\label{fig:fig3}
\end{figure}

@@

We re-use the above queried ETOX-IDs to get further information from ETOX, namely the MAC-EQS:
<<search_targests, message=FALSE, warning=FALSE>>=
eqs <- etox_targets(ids$etoxid)
ids$mac <- sapply(eqs, function(y){
  if (length(y) == 1 && is.na(y)) {
    return(NA) 
  } else {
    res <- y$res
    min(res[res$Country_or_Region == 'EEC / EU' & 
              res$Designation == 'MAC-EQS', 'Value_Target_LR'])
  }
})
@

@@

<<mac>>=
(mac <- with(ids, ids[!is.na(mac) & is.finite(mac), 
                      c('etoxid', 'query', 'mac')]))
@

The \code{get_etoxid()} function used to search ETOX-IDs returns also the original substance name (\code{query}),
so that we can easily join the table with MAC values with the measurements table :
<<>>=
jagst_eqs <- merge(jagst, mac, by.x = 'substance', by.y = 'query')
head(jagst_eqs)
@

@@

<<check_format>>=
is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
is.cas('64-17-6')
@

or web resources like ChemSpider
<<check_web>>=
is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
  type = 'chemspider')
@

>>>>

\documentclass[article, shortnames]{jss}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{Sweave}

\usepackage[utf8]{inputenc} %unicode support
\usepackage{booktabs}
\usepackage{fixltx2e}

@@

% citation alias for LUBW institution
\defcitealias{lubw_2016}{LUBW (2016)}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("webchem")
R> library("webchem")
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("jagst")
R> head(jagst)
\end{CodeInput}
\begin{CodeOutput}
        date          substance value qual
1 2013-01-04 2,4-Dimethylphenol 0.006    <
2 2013-01-29 2,4-Dimethylphenol 0.006    <
3 2013-02-26 2,4-Dimethylphenol 0.006    <
4 2013-03-26 2,4-Dimethylphenol 0.006    <
5 2013-04-23 2,4-Dimethylphenol 0.006    <
6 2013-05-22 2,4-Dimethylphenol 0.006    <
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("lc50")
R> head(lc50)
\end{CodeInput}
\begin{CodeOutput}
       cas        value
4  50-29-3    12.415277
12 52-68-6     1.282980
15 55-38-9    12.168138
18 56-23-5 35000.000000
21 56-38-2     1.539119
36 57-74-9    98.400000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> subs <- unique(jagst$substance)
R> ids <- get_etoxid(subs, match = 'best')
R> head(ids)
\end{CodeInput}
\begin{CodeOutput}
  etoxid                           match distance                  query
1   8668     2,4-Dimethylphenol ( 8668 )        0     2,4-Dimethylphenol
2   8494 4-Chlor-2-methylphenol ( 8494 )        0 4-Chlor-2-methylphenol
3   <NA>                            <NA>     <NA>     4-para-nonylphenol
4   8397                Atrazin ( 8397 )        0                Atrazin
5   7240                 Benzol ( 7240 )        0                 Benzol
6   7331        Desethylatrazin ( 7331 )        0        Desethylatrazin
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> etox_data <- etox_basic(ids$etoxid)
\end{CodeInput}
\end{CodeChunk}

\pkg{webchem} always returns a named list (one entry for each substance) and the available information content can be very voluminous.
Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
\begin{CodeChunk}
\begin{CodeInput}
R> etox_cas <- cas(etox_data)
R> head(etox_cas)
\end{CodeInput}
\begin{CodeOutput}
       8668        8494        <NA>        8397        7240        7331 
 "105-67-9" "1570-64-5"          NA "1912-24-9"   "71-43-2" "6190-65-4" 
\end{CodeOutput}
\end{CodeChunk}

@@

In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
\begin{CodeChunk}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> cids <- get_cid(etox_cas)
R> pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
R> pc_smiles <- smiles(pc_data)
\end{CodeInput}
\end{CodeChunk}

or ChemSpider
\begin{CodeChunk}
\begin{CodeInput}
R> csids <- get_csid(etox_cas, token = token)
R> cs_data <- cs_compinfo(csids, token = token)
R> cs_inchikey <- inchikey(cs_data)
\end{CodeInput}
\end{CodeChunk}

Finally, we combine the queried data into one data.frame
\begin{CodeChunk}
\begin{CodeInput}
R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+    cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
+    stringsAsFactors = FALSE)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{table}[ht]
\centering
\begin{CodeChunk}
% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Tue Sep 20 16:11:04 2016
\begin{tabular}{llllll}
  \toprule
Name & CAS & SMILES & CID & InChIKey & CSID \\ 
  \midrule
2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & - & - \\ 
  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & - & - \\ 
  4-para-nonylphenol & - & - & - & - & - \\ 
  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & - & - \\ 
  Benzol & 71-43-2 & C1=CC=C... & 241 & - & - \\ 
  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & - & - \\ 
   \bottomrule
\end{tabular}
\end{CodeChunk}
\caption{Identifiers for the jagst data sets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
\label{tab:comptable}
\end{table}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> aw_data <- aw_query(lc50$cas, type = 'cas')
\end{CodeInput}
\begin{CodeOutput}
Warning: closing unused connection 35 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=67-66-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 34 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=75-25-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 33 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=108-88-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 32 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=127-18-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 31 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=5915-41-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 30 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=886-50-0&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 29 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=122-34-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 28 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=60207-90-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 27 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=139-40-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 26 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=23103-98-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 25 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1763-23-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 24 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=335-67-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 23 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=307-24-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 22 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=375-22-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 21 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=40487-42-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 20 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=66246-88-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 19 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=51218-45-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 18 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=67129-08-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 17 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=57837-19-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 16 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=94-74-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 15 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=34123-59-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 14 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=28159-98-0&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 13 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=83164-33-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 12 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=134-62-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 11 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=75-09-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 10 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=30125-63-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 9 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=6190-65-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 8 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=71-43-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 7 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1912-24-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 6 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1570-64-5&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\begin{CodeOutput}
Warning: closing unused connection 5 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=105-67-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> igroup <- sapply(aw_data, function(y) y$subactivity[1])
R> igroup[1:3]
\end{CodeInput}
\begin{CodeOutput}
                                  50-29-3 
            "organochlorine insecticides" 
                                  52-68-6 
               "phosphonate insecticides" 
                                  55-38-9 
"phenyl organothiophosphate insecticides" 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{figure}[ht]
\begin{CodeChunk}

{\centering \includegraphics[width=0.7\textwidth]{plot_lc50-1}

}

\end{CodeChunk}
\caption{Toxicity of different pesticide groups. LC\textsubscript{50} values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
\label{fig:fig2}
\end{figure}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pp_data <- pp_query(lc50$cas)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> lc50$logp <- sapply(pp_data, function(y) {
+    if (length(y) == 1 && is.na(y))
+      return(NA)
+    y$prop$value[y$prop$variable == 'Log P (octanol-water)']
+  })
\end{CodeInput}
\end{CodeChunk}

@@

\begin{figure}[ht]
\begin{CodeChunk}

{\centering \includegraphics[width=0.7\textwidth]{plot_qsar-1}

}

\end{CodeChunk}
\caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
Blue line indicates the regression model ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log} P$, RMSE = 1.45).}
\label{fig:fig3}
\end{figure}

@@

We re-use the above queried ETOX-IDs to get further information from ETOX, namely the MAC-EQS:
\begin{CodeChunk}
\begin{CodeInput}
R> eqs <- etox_targets(ids$etoxid)
R> ids$mac <- sapply(eqs, function(y){
+    if (length(y) == 1 && is.na(y)) {
+      return(NA) 
+    } else {
+      res <- y$res
+      min(res[res$Country_or_Region == 'EEC / EU' & 
+                res$Designation == 'MAC-EQS', 'Value_Target_LR'])
+    }
+  })
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> (mac <- with(ids, ids[!is.na(mac) & is.finite(mac), 
+                        c('etoxid', 'query', 'mac')]))
\end{CodeInput}
\begin{CodeOutput}
   etoxid       query    mac
4    8397     Atrazin  2.000
5    7240      Benzol 50.000
11   8836     Irgarol  0.016
12   7442 Isoproturon  1.000
29   8756   Terbutryn  0.034
\end{CodeOutput}
\end{CodeChunk}

The \code{get_etoxid()} function used to search ETOX-IDs returns also the original substance name (\code{query}),
so that we can easily join the table with MAC values with the measurements table :
\begin{CodeChunk}
\begin{CodeInput}
R> jagst_eqs <- merge(jagst, mac, by.x = 'substance', by.y = 'query')
R> head(jagst_eqs)
\end{CodeInput}
\begin{CodeOutput}
  substance       date  value qual etoxid mac
1   Atrazin 2013-09-10 0.0068    =   8397   2
2   Atrazin 2013-10-08 0.0072    =   8397   2
3   Atrazin 2013-03-26 0.0040    =   8397   2
4   Atrazin 2013-04-23 0.0048    =   8397   2
5   Atrazin 2013-11-05 0.0036    =   8397   2
6   Atrazin 2013-07-16 0.0052    =   8397   2
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
\end{CodeInput}
\begin{CodeOutput}
Hyphens not at position 15 and 26.
\end{CodeOutput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\begin{CodeInput}
R> is.cas('64-17-6')
\end{CodeInput}
\begin{CodeOutput}
Checksum is not correct! 5 vs. 6
\end{CodeOutput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\end{CodeChunk}

or web resources like ChemSpider
\begin{CodeChunk}
\begin{CodeInput}
R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
+    type = 'chemspider')
\end{CodeInput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\end{CodeChunk}
